Recent demography drives changes in linked selection across the maize genome

### 2015-11-11 manuscript.tex
edits to intro., results
tweak

In addition to our simple bottleneck model, we investigated two alternative approaches for demographic inference.  First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion. 
This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
Finally, we applied a model-free coalescent approach \cite{schiffels2014} using a subset of our samples. 
Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).

@@

Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because diversity in such cases  is not necessarily reduced \cite{innan2004,messer2013}. 
Such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes. 
To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% quantile of H12 (Figure \ref{sFig:H12}A).
%\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
%TB Says: Sites in the top 20% of H12 are unlikely to be under background selection, so maize vs teo background selection pattern at these sites seems strange to investigate.
We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.

>>>>

In addition to our simple bottleneck model, we investigated two alternative approaches for demographic inference.  First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion. 
This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
Finally, we applied a model-free coalescent approach \cite{schiffels2014} using a subset of our samples. 
Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
Our assessment of the historical demography of maize and teosinte provides important context for subsequent analyses of linked selection.

@@

Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity around the substitution may be reduced to a lesser degree \cite{innan2004,messer2013}. 
Nonetheless, such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity at linked sites in genes. 
To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% quantile of H12 (Figure \ref{sFig:H12}A).
%\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
%TB Says: Sites in the top 20% of H12 are unlikely to be under background selection, so maize vs teo background selection pattern at these sites seems strange to investigate.
We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
